Thermodynamic fragments for magic states in quantum computation

### 2021-04-01 magic_paper.tex
Cleaned up Sec.III

Within the Wigner representation, all negatively represented states are magic states, although in general the converse is not true~\cite{cit:campbell}. However, it is known that all positively represented states used in Clifford circuits admit an efficient classical description, so it is natural to focus on the negatively represented magic states. The free states in any magic theory must therefore lie in the set
\begin{equation}
    \Fmax \coloneqq \{ \rho: \W[\bmx]{\rho} \geq 0 \text{ for all } \bmx \in \cal{P}_d\}.
\end{equation}
Our focus is on states with negativity, so the particular choice of free states is not critical for our analysis. The remaining component that defines any magic theory is the set of allowed (free) quantum operations, i.e. the ones that are considered free. The assumption we require on free operations is simply that they send any free state to another free state,meaning that $\F$ must be closed, which is a common assumption among all magic theories.

\nick{``the condition that $\E$ does not generate Wigner negativity is an essential requirement of any magic theory'': this is not true for stabilizer-preserving operations (SPOs)~\cite{cit:ahmadi}. Note that ``completely'' SPOs~\cite{cit:seddon} are a strict subset. However, SPOs are a subset of Wang's CPWP operations~\cite{Wang_2019} so our results hold for SPOs too.}

@@

Our formulation will not consider model-specific limitations, but will still inject broad physical constraints. We proceed by considering imposing equilibrium, or fixed-point, restrictions on the actual operations that are accessible to us. While such restrictions could be viewed as the existence of some non-trivial thermodynamic constraint, they could equally be viewed as encoding a high noise-bias or difficulty in realising a particular quantum gate. We can therefore define the following sub-theory of any magic theory $\R$.
\begin{definition}[\textbf{$\boldsymbol\sigma$--fragment}]\label{def:sigmafrag}
   Given a theory of magic $\R = (\F, \O)$, the \emph{$\sigma$--fragment of $\R$} is the sub-theory $\R_\sigma = (\F, \O_\sigma)$, in which the free operations are 
   \begin{equation}
        \O_\sigma \coloneqq \{ \E \in \O: \E(\sigma) = \sigma \},
    \end{equation}
namely those operations that leave the state $\sigma \in \F$ invariant.
\end{definition}
This notion provides a simple way to break up any theory into smaller, more manageable parts that are destinguished by additional physical constraints. Moreover, the union over all fragments returns the parent theory, so we are not discarding any information by breaking up the theory in this way. We make this statement precise as follows.
\begin{theorem}\label{thm:frag}
    Let $\R = (\F, \O)$ be a theory of magic.
Then a transformation $\rho \longrightarrow \tau$ is possible in $\R$ if and only if the transformation $\rho \longrightarrow \tau$ is possible in at least one $\sigma$--fragment of $\R$.
\end{theorem}
The proof of this is straightforward.
\begin{proof}
    Suppose the interconversion is possible in a $\sigma$--fragment via some $\E \in \O_\sigma$. But since $\O_\sigma \subseteq \O$ it is also possible in $\R$. Conversely, suppose the transformation is possible in $\R$ via some $\E$ in $\O$. The free states $\F$ are a closed, convex set and moreover the image of $\F$ under the map $\E$ is in $\F$. By the Brouwer Fixed-Point theorem~\cite{cit:brouwer} this mapping must therefore have a fixed point $\sigma \in \F$. Therefore $\E \in \O_\sigma$ and so the interconversion is possible in this fragment.
\end{proof}

We can now consider magic distillation rates within a particular $\sigma$--fragment and then see how the bounds on the distillation rates depend on the particular fragment we work within. Therefore, this route can potentially shed light on how the physics of a protocol constrain distillation rates above and beyond global constraints from the parent resource theory. For example, how does temperature affect distillation bounds? Do isotropic maps have better bounds, or does bias towards a pure stabilizer state fixed point provide more freedom?
\nick{We probably want to provide clear answers to these later on}

\begin{figure}[t]
    \centering
        \includegraphics[scale=0.3
        ]{figs/operations.pdf}
    \caption{\textbf{Decomposition of a magic theory $\R$ intro $\sigma$--fragments.} 
	Examples of magic theories ($\so$: Stabilizer operations, $\Omax$: Completely positive-Wigner-preserving operations, $\rcu$: Random Clifford Unitaries -- subclass of $\so$) involve operations denoted by the two yellow regions, with other established magic theories  between them.
    We introduce $\sigma$--fragments $\O_\sigma$ defined for all free states $\sigma$ that cover $\Omax$. 
    Each $\O_\sigma$ is extensible to a set of stochastic maps outside the $\cptp$ operations.
    Within each $\sigma$--fragment $\bmd$--majorisation can be used allowing for tractable approach towards the study of magic state interconversion.
    }
    \label{fig:zoo}
\end{figure}

@@

We can now show that each magic fragment admits constraints that do not take the form of state monotones. The development of any monotone $\M(\rho)$ for a magic theory $\R$ can be viewed abstractly as
\begin{equation}
\M : \mbox{(pre-order of states)} \mapsto \mbox{(total order of numerical values).} \nonumber
\end{equation}
More precisely, within the theory we order magic states as $\rho_1 \succ \rho_2$ whenever there is an operation $\E \in \O$ such that $\E(\rho_1) = \rho_2$. However, in general many states will not be interconvertible in either direction, and so this ordering is a pre-order in the sense that incomparable states exist. A magic monotone projects this pre-order of states onto the number line, and so through this lens any two states are directly comparable in terms of the values of $\M(\rho_1)$ and $\M(\rho_2)$. Of course, the price is that the monotone provides only a necessary, but not sufficient, condition for the transformation $\rho_1 \rightarrow \rho_2$.

The restriction of a magic theory $\R$ to some fragment $\R_\sigma$ can now be interpreted as a generalisation of the magic monotone construction: instead of mapping the pre-order of states in $\R$ onto the number line, we instead map this pre-order onto \emph{a simpler pre-order of states.} Therefore, we are retaining more of the original parent theory, while working in a scenario that admits greater structure and simplicity. We can view this as a mapping $\P_\sigma: \R \rightarrow \R_\sigma$ that restricts to a sub-theory (a $\sigma$--fragment) of the parent theory $\R$. Thus,
\begin{equation}
\P_\sigma: \mbox{(pre-order)} \longrightarrow \mbox{(simpler pre-order).}
\end{equation}
By framing the projection onto a fragment in this light, we can view Theorem \ref{thm:frag} as a statement that the collection $\{\P_\sigma\}$ provides a complete set of generalized monotones for the theory.

The question is now whether the sub-theories can be analysed to provide additional insights into magic distillation protocols. We shall do this by working in the Wigner representation of a given $\sigma$--fragment. The problem of magic state distillation then becomes that of processing negativity under stochastic maps $W_{\E}( \y |\x)$ that leave a particular quasi-distribution $W_\sigma(\x)$ invariant. However $\sigma \in \F$ is is positively represented and so the sub-theory has an invariant classical probability distribution as a fixed point. This structure allows us to make use of a range of classical tools, and in particular majorization theory.

Majorisation~\cite{cit:marshall} is a collection of powerful tools that has recently found many applications in quantum information theory~\cite{Nielsen_1999, cit:cwiklinski, cit:lostaglio2, cit:gour, cit:gour2, Horodecki_2003, Vallejos_2021}.
It describes the disorder of distributions that undergo stochastic transformations. In its simplest form it defines a pre-order on probability distributions. Given two distributions $\p= (p_1, \dots, p_n)$ and $\q = (q_1, \dots, q_n)$ over $n$ outcomes, we say that $\p$ majorises $\q$, denoted $\p \succ \q$, if there is a bistochastic map $A=(A_{ij})$ such that $A\p = \q$. Here, bistochastic means that $\sum_i A_{ij} = \sum_j A_{ij} = 1$. A well known result tells us that the condition $ \p \succ \q$ over probability distributions is equivalent to $n-1$ inequalities, and these conditions are easily checked.

However, there is a natural generalisation of this, called \emph{relative majorization} (also known as $d$--majorization, or in the context of thermodynamics, thermo-majorization) that dates back to the 1950s. For a fixed probability distribution $\r = (r_1, \dots, r_n)$ we define majorization relative to $\r$ as follows:
\begin{equation}
\p \succ_{\r} \q,
\end{equation}
if and only if there exists a stochastic map $A$ such that $A\r = \r$ and $A \p = \q$. In order words, the relative majorization pre-order coincides with the image of $\p$ under stochastic maps that have $\r$ as a fixed point. It is readily checked that the original majorization condition between probability distributions corresponds to the case $\r = (1/n, \dots, 1/n)$. The pre-order that results from relative majorization is also easily checked and again corresponds to checking a small number of inequalities.

However, it turns out that relative majorization is equally applicable to \emph{quasi-probability} distributions, and therefore can be applied to the Wigner representation of a given fragment. To our knowledge, this application of majorization to quasi-distributions in  physics has not been considered previously. The caveat here is that relative majorization ranges over all stochastic maps with a given fixed point, while in the case of the Wigner representation there is an additional restriction that the symplectic structure is respected. This implies that any relative majorization conditions provide necessary, but not sufficient, conditions on magic state interconversions in the fragment. However, this observation raises a question that to our knowledge has not been previously considered in the classical statistical mechanics literature: what are the conditions for symplectic majorization of classical statistical mechanics? We shall return to this question later on when we discuss how our approach could be used to construct concrete lower-bounds on distillation rates.

@@

Given a vector $\x$ we define $\x^\downarrow$ to be the re-arrangement of the components of $\x$ into decreasing order. Now, given two $n$--component vectors $\w$ and $\r$ we define the \emph{Lorenz curve} of $\w$ relative to $\r$, denoted $L_{\w|\r}(x)$, as the piece-wise linear function that passes through $(0,0)$ and the $n$ points
\begin{equation}\label{eq:lorenz}
        (x_k,L_{\w|\r}(x_k)) =\left( \frac{1}{R}\sum_{i=1}^k r_{\pi(i)}, \sum_{i=1}^k w_{\pi(i)} \right),
\end{equation}
where $R:= \sum_{i=1}^n r_i$, and $\pi$ is the permutation on $n$ objects that sends $\tilde{\w}$ to $\tilde{\w}^{\downarrow}$, where $\tilde{w}_i := w_i/r_i$ is the vector of component-wise ratios of $\w$ and $\r$. The form of this requires that $\r$ has no zero components, which we shall assume without loss of generality as in any physical situation the rank of a quantum state is not an operationally meaningful quantity.

If $\w$ and $\r$ are both probability distributions then the Lorenz curve is defined on the interval $[0,1]$, and rises monotonically until it reaches the value $1$ at $x=1$. The value $L_{\w|\r}(x) = 1$ is a global upper bound to the function, only attained at the end-point $x=1$. However, if $\w$ is a \emph{quasi-probability} distribution with negative values, and $\r$ a regular probability distribution things are different. Now the Lorenz curve is no longer monotone increasing, but is a concave function that breaks through the $L_{\w|\r}(x) = 1$ barrier at an interior point and attains some non-trivial maximum $L_\star$ above the value $1$, before decreasing monotonically to $L_{\w|\r}(x)= 1$ at the end-point $x=1$. See Figure \ref{fig:lcs} for examples of non-monotonic Lorenz curves for quasi-distributions. For quasi-distributions within quantum theory, the breaking of this barrier is associated with the degree of non-classicality. Since majorization provides a measure of order/disorder this is consistent with the intuition that negativity, for example within the context of entanglement theory or quantum computing, can provide a greater form of order than is possible within classical theory.

\begin{figure}
    \centering
    \includegraphics[height=4.5cm]{figs/lc_strange.pdf}
    \caption{\textbf{A `heretical' family of Lorenz curves.} Traditionally, Lorenz curves for distributions are monotone increasing cumulant functions that reach a maximum value of $1$. In contrast, Lorenz curves for magic states break through the value of $L(x)=1$ due to the presence of negativity in the associated quasi-probability distributions. The above family of curves correspond to multiple copies of noisy Strange states $\rho_{\rm{S}}(\epsilon)^{\otimes n}$ for $n=1,2,3,4$ within a particular stabilizer fragment. Solid lines represent pure Strange states, while dashed lines represent noisy Strange states with depolarising noise $\epsilon = 0.1$.
    }
    \label{fig:lcs}
\end{figure}

Given the definition of Lorenz curves, the condition for relative majorization can now be compactly phrased. We have that $\w_1 \succ_{\r} \w_2$ if and only if $L_{\w_1 | \r} (x) \ge L_{\w_2 | \r}(x)$ for all $x \in [0,1]$. In other words, the Lorenz curve of $\w_1$ relative to $\r$ is never below the Lorenz curve of $\w_2$ relative to $\r$. Phrased more abstractly, the relative majorization pre-order corresponds to the standard ordering of concave, piece-wise linear functions passing through $(0,0)$ and $(1,1)$. Another neat geometric formulation of relative majorization exists that simply states that the $L_1$-norm angle between $\w_1$ and $\r$ is greater than or equal to the $L_1$-norm angle between $\w_2$ and $\r$ (see \ddd{[CITE]} for details).

@@

If we consider some $\sigma \in \F$, and its corresponding fragment $\R_\sigma$, then we are typically interested in the ability to transform many copies of some magic state $\rho$ into more pure forms of magic. The state $\rho$ is assumed to have negativity in the Wigner representation, and so $W_\rho(\x) < 0$ for some regions of $\x \in \P_d$. We again note that we restrict to odd-dimensional quantum systems, for simplicity. In contrast the state $\sigma$ will have a Wigner distribution $W_\sigma(\x)$ that forms a genuine probability distribution. On the technical assumption that $\sigma$ is full rank, then we also have that $W_\sigma(\x)$ is non-negative on the discrete phase space $\P_d$. Any non-full rank states can be handled as a limiting case in which we mix in a small fraction $\epsilon (\I/d)$ to the quantum state $\sigma$, and later take $\epsilon$ tending to zero.

As discussed already, the free operations within the magic theory $\R$ are represented by stochastic maps, and within $\R_\sigma$ by stochastic maps that leave $W_\sigma(\x)$ invariant. Therefore a necessary condition for magic state transformations $\rho_1 \rightarrow \rho_2$ within $\R_\sigma$ will be that 
\begin{equation}
W_{\rho_1} \succ_{W_{\sigma}} W_{\rho_2},
\end{equation}
or put another way, relative to $W_\sigma$ the quasi-distribution $W_{\rho_1}$ is more ordered than the quasi-distribution $W_{\rho_2}$.

This in turn can be expressed in terms of Lorenz curves. To simplify notation we denote by $L_{\rho | \sigma}(x)$ the Lorenz curve $L_{W_{\rho} | W_{\sigma}} (x)$. We therefore have that
\begin{equation}
\rho_1 \rightarrow \rho_2 \mbox{ within }\R_\sigma \mbox{ implies that } L_{\rho_1 |\sigma} (x) \ge L_{\rho_2 |\sigma} (x),
\end{equation}
for all $x \in [0,1]$. Therefore, the constraint in terms of the Lorenz curves for the magic states can be used to impose upper bounds on magic distillation rates. Note however, that this is not a single numerical constraint, but is a \emph{family} of constraints. Indeed for $n$ copies of a $d$ dimensional system the number of terms in $W_{\rho}$ is $d^{2n}$ and so the Lorenz curve condition corresponds to \emph{exponentially} many constraints.

%We can approach any magic theory through a thermodynamic lens, and in doing so we are provided with valuable insights on the structure of the theory. 
%Firstly, any free state, for example a stabiliser state, can be viewed as a thermal state $\gamma_\beta$.
%Without loss of generality we can always write a quantum state $\sigma$ as a thermal state, $\sigma = \gamma_\beta \coloneqq \frac{1}{\Z_\beta} e^{-\beta H}$ for some $\beta \geq 0$ and Hamiltonian $H$ (either effective or actual)\footnote{Technicalities arise for the case where $\sigma$ is not full rank, but this can be still described via $ \beta \rightarrow 0$ limiting process.}.
%We can also view the set of free operations $\O_\sigma$ as a counterpart of thermal operations, in the sense that any operation $\E$ in $\O_\sigma$ preserves state $\sigma$. 
%
%It is then apparent that the pre-order $\prec_{\R_\sigma}$ between the operations in the $\sigma$--fragment follows the rules of $\bmd$--majorisation as outlined in~\cref{sec:major}.
%In particular, the pre-order $\prec_{\R_\sigma}$ of the $\sigma$--fragment $\R_\sigma = (\F, \O_\sigma)$ between $d$--dimensional states corresponds to the majorisation pre-order $\prec_{\W{\sigma}}$ between their $d^2$--dimensional Wigner distributions.
%For simplicity we shall merge the notation into $\prec_\sigma$, as there is little risk of confusion.
%
%Note that the Wigner components of an $n$--copy state $\rho^{\otimes n}$ can be calculated directly from $\W{\rho}$ by convolution of the distribution with itself,
%\begin{equation}
%	\W{\rho^{\otimes n}} = \W{\rho}^{\otimes n},
%\end{equation}
%where $\otimes$ can be interpreted as the usual Kronecker product in the last expression.
%We may use the vector notation $\bmw(\rho) = (w(\rho)_i)_{i=1,\dots,d^2}$ for the Wigner distribution, in which case we can express the Kronecker product as
%\begin{equation}
%	\W{\rho} \otimes \W{\rho} = (w(\rho)_i w(\rho)_j)_{i,j=1,\dots,d^2}.
%\end{equation}
%The correspondence between the phase space and pure vector representations of the Wigner distribution is discussed more in~\cref{app:cmpairs}.
%The vector notation will be useful in the proof of our main result,~\cref{thm:stab_bounds}
%
%Furthermore, we restrict our analysis to $\sigma$--fragments, where $\sigma$ is full-rank.
%This is justified as majorisation is continuous between fragments, in the sense that the pre-orders $\prec_\sigma$ and $\prec_{\sigma'}$ are equivalent for states $\sigma$ and $\sigma'$ which are $\epsilon$--close. \nick{CHECK}
%Therefore, majorisation analysis is robust under imperfections in the experimental implementation of quantum operations.
%
%In particular, we can always use common noise effects to approximate any $\sigma$--fragment where $\sigma$ is not full-rank by the $\sigma'$--fragment where $\sigma'$ is a noisy approximation of $\sigma$. 
%For example, inducing depolarising noise, we can write $\sigma' = (1-\epsilon)\sigma + \epsilon\frac{1}{d}\id$, for some infinitesimal $\epsilon > 0$, so that $\sigma'$ is full-rank and arbitrarily close to $\sigma$.
%Important examples of such $\sigma$--fragments include pure stabiliser states which are rank-$1$, e.g. the zero state depicted in~\cref{fig:zero}. 
%Operations in such fragments include important stabiliser operations like the replacement channel, $\E(\rho) = \sigma$ for all states $\rho$.
%
%\begin{theorem}\label{thm:sigmamajor}
%    Let $\R = (\F, \O)$ be a theory of magic. If $\rho \longrightarrow \tau$  in $\R$ then  $\W{\tau} \prec_\sigma \W{\rho}$ within at least one $\sigma$--fragment.
%\end{theorem}
%\begin{proof}
%Suppose we can convert $\rho$ into $\tau$ in the magic theory. 
%Thus there is some $\O_\sigma$ and some $\E \in \O_\sigma$ such that $\E(\rho) = \tau$, and $\E(\sigma)=\sigma$. 
%Therefore, the Wigner distribution of this free operation satisfies $\W{\E} \in \stochw$ and $\W{\E}\W{\rho} = \W{\tau}$. 
%Since $\sigma$ is full-rank and free, its Wigner distribution is strictly positive in all components, so it directly follows from~\cref{def:dmajor} that $\W{\tau} \prec_{\sigma} \W{\rho}$.
%\end{proof}
%The converse is in general not true, since stochastic matrices do not necessarily correspond to valid quantum operations.
%
%The result of~\cref{thm:sigmamajor} can be understood as an extension of the idea of a magic monotone, where we replace $\M(\tau) \leq \M(\rho)$ with $\W{\tau} \prec_\sigma \W{\rho}$. 
%The physical difference between the two expressions is that the majorisation ordering occurs in a specific $\sigma$--fragment.
%Therefore, majorisation constraints can be used to place upper bounds on magic state distillation in a way that allows one to incorporate the physics of the allowed operations -- specifically, it enables us to bound how much magic can be distilled via quantum operations that, for example, preserve the equilibrium state of the system, or via operations that are symmetric about the $Z$-axis of the Bloch sphere.
%We discuss distillation upper bounds in detail in~\cref{sec:unital,sec:stab}

>>>>

Within the Wigner representation, all negatively represented states are magic states, although in general the converse is not true~\cite{cit:campbell}. However, it is known that all positively represented states used in Clifford circuits admit an efficient classical description, so it is natural to focus on the negatively represented magic states. The free states in any magic theory must therefore lie in the set
\begin{equation}
    \Fmax \coloneqq \{ \rho: \W[\bmx]{\rho} \geq 0 \text{ for all } \bmx \in \cal{P}_d\}.
\end{equation}
Our focus is on states with negativity, so the particular choice of free states is not critical for our analysis. The remaining component that defines any magic theory is the set of allowed (free) quantum operations, i.e. the ones that are considered free. The assumption we require on free operations is simply that they send any free state to another free state, meaning that a resource state cannot be created freely, a common assumption among all magic theories.

\nick{``the condition that $\id \otimes \E$ does not generate Wigner negativity is an essential requirement of any magic theory'': this is not true for stabilizer-preserving operations (SPOs)~\cite{cit:ahmadi}. Note that ``completely'' SPOs~\cite{cit:seddon} are a strict subset. However, SPOs are a subset of Wang's CPWP operations~\cite{Wang_2019} so our results hold for SPOs too.}

@@

Our formulation will not consider model-specific limitations, but will still inject broad physical constraints. We proceed by considering fixed-point, or ``equilibrium'', restrictions on the actual operations that are accessible to us. While such restrictions could be viewed as the existence of some non-trivial thermodynamic constraint, they could equally be viewed as encoding a high noise-bias or difficulty in realising a particular quantum gate. We can therefore define the following sub-theory of any magic theory $\R$.
\begin{definition}[\textbf{$\boldsymbol\sigma$--fragment}]\label{def:sigmafrag}
   Given a theory of magic $\R = (\F, \O)$, the \emph{$\sigma$--fragment of $\R$} is the sub-theory $\R_\sigma = (\F, \O_\sigma)$, in which the free operations are 
   \begin{equation}
        \O_\sigma \coloneqq \{ \E \in \O: \E(\sigma) = \sigma \},
    \end{equation}
namely those operations that leave the state $\sigma \in \F$ invariant.
\end{definition}
This notion provides a simple way to break up any theory into smaller, more manageable parts that are destinguished by additional physical constraints. Moreover, the union over all fragments returns the parent theory, so we are not discarding any information by breaking up the theory in this way. We make this statement precise as follows.
\begin{theorem}\label{thm:frag}
    Let $\R = (\F, \O)$ be a theory of magic.
Then a transformation $\rho \longrightarrow \tau$ is possible in $\R$ if and only if the transformation $\rho \longrightarrow \tau$ is possible in at least one $\sigma$--fragment of $\R$.
\end{theorem}
The proof of this is straightforward.
\begin{proof}
    Suppose the interconversion is possible in a $\sigma$--fragment via some $\E \in \O_\sigma$. But since $\O_\sigma \subseteq \O$ it is also possible in $\R$. Conversely, suppose the transformation is possible in $\R$ via some $\E$ in $\O$. The free states $\F$ are a closed, bounded set and moreover the image of $\F$ under the map $\E$ is in $\F$. By the Brouwer Fixed-Point theorem~\cite{cit:brouwer} this mapping must therefore have a fixed point $\sigma \in \F$. Therefore $\E \in \O_\sigma$ and so the interconversion is possible in this fragment.
\end{proof}

We can now consider magic distillation rates within $\sigma$--fragments and then see how the bounds on the distillation rates depend on the choice of the fragment we work within. Therefore, this route can potentially shed light on how the physics of a protocol constrain distillation rates above and beyond global constraints from the parent resource theory. For example, how does temperature affect distillation bounds? Do isotropic maps have better bounds, or does bias towards some pure stabilizer state provide more freedom?
\nick{remember to re-address these questions after deriving our bounds later on}

\begin{figure}[t]
    \centering
        \includegraphics[scale=0.3
        ]{figs/operations.pdf}
    \caption{\textbf{Decomposition of a magic theory $\R$ intro $\sigma$--fragments.} 
	Examples of magic theories ($\so$: Stabilizer operations, $\Omax$: Completely positive-Wigner-preserving operations, $\rcu$: Random Clifford Unitaries -- subclass of $\so$) are labelled, with other established magic theories also contained within the yellow regions.
    We introduce $\sigma$--fragments $\O_\sigma$ defined for all free states $\sigma$ that cover $\Omax$. 
    Each $\O_\sigma$ is extensible to a set of stochastic maps outside the $\cptp$ operations.
    Within each $\sigma$--fragment, $\bmd$--majorisation can be used allowing for a tractable approach towards the study of magic state interconversions.
    }
    \label{fig:zoo}
\end{figure}

@@

We can now show that each magic fragment admits constraints that do not take the form of state monotones. 
The development of any monotone $\M(\rho)$ for a magic theory $\R$ can be viewed abstractly as
\begin{equation}
\M : \mbox{(pre-order of states)} \mapsto \mbox{(total order of numbers).} \nonumber
\end{equation}
More precisely, within the theory we order magic states as $\rho_1 \succ \rho_2$ whenever there is an operation $\E \in \O$ such that $\E(\rho_1) = \rho_2$. However, in general many states will not be interconvertible in either direction, and so this ordering is a pre-order in the sense that incomparable states exist. A magic monotone projects this pre-order of states onto the real number line, and, through this lens, any two states are directly comparable in terms of their mapped values $\M(\rho_1)$, $\M(\rho_2)$. Of course, the price is that the monotone provides only a necessary, but not sufficient, condition for the transformation $\rho_1 \rightarrow \rho_2$.

The restriction of a magic theory $\R$ to some fragment $\R_\sigma$ can now be interpreted as a generalisation of the magic monotone construction: instead of mapping the pre-order of states in $\R$ onto the number line, we instead map this pre-order onto \emph{a simpler pre-order of states}. Therefore, we are retaining more of the original parent theory, while working in a scenario that admits greater structure and simplicity. We can view this as a mapping $\P_\sigma: \R \rightarrow \R_\sigma$ that restricts to a sub-theory (a $\sigma$--fragment) of the parent theory $\R$. Thus,
\begin{equation}
\P_\sigma: \mbox{(pre-order)} \longrightarrow \mbox{(simpler pre-order).}
\end{equation}
By framing the projection onto a fragment in this light, we can view Theorem \ref{thm:frag} as a statement that the collection $\{\P_\sigma\}_{\sigma \in \F}$ provides a complete set of generalized monotones for the theory.

The question is now whether the sub-theories can be analysed to provide additional insights into magic distillation protocols. We shall provide an answer by working in the Wigner representation of a given $\sigma$--fragment. The problem of magic state distillation then becomes that of processing negativity under stochastic maps $W_{\E}( \y |\x)$ that leave a particular quasi-distribution $W_\sigma(\x)$ invariant. However, $\sigma \in \F$ is positively represented, so the sub-theory has an invariant classical probability distribution as a fixed point. This structure allows us to make use of a range of classical tools, and in particular majorization theory.

Majorisation~\cite{cit:marshall} is a collection of powerful tools that has recently found many applications in quantum information theory~\cite{Nielsen_1999, cit:cwiklinski, cit:lostaglio2, cit:gour, cit:gour2, Horodecki_2003, Vallejos_2021}.
It describes the disorder of distributions that undergo stochastic transformations. In its simplest form it defines a pre-order on probability distributions. Given two distributions $\p= (p_1, \dots, p_n)$ and $\q = (q_1, \dots, q_n)$ over $n$ outcomes, we say that $\p$ majorises $\q$, denoted $\p \succ \q$, if there exists a bistochastic map $A = (A_{ij})$ such that $A\p = \q$, where bistochastic means that $A_{ij} \geq 0$ and $\sum_i A_{ij} = \sum_j A_{ij} = 1$. A well-known result~\cite{cit:marshall} tells us that the condition $ \p \succ \q$ over probability distributions is equivalent to $n-1$ inequalities, which can be easily checked.

There is a natural generalisation, which we call \emph{relative majorization} (also known as $d$--majorization~\cite{Veinott_1971}, or in the context of thermodynamics, thermo-majorization)\footnote{The term ``relative majorization'' is often used~\cite{Ruch_1976} to describe a further generalisation of $\bmd$-majorization, but here we use it as equivalent.}. For a fixed probability distribution $\r = (r_1, \dots, r_n)$ with positive components, we define majorization relative to $\r$ as follows:
\begin{equation}
\p \succ_{\r} \q,
\end{equation}
if and only if there exists a stochastic map $A$ such that $A\r = \r$ and $A \p = \q$. In order words, the relative majorization pre-order coincides with the image of $\p$ under stochastic maps that have $\r$ as a fixed point. It is readily checked that the original majorization condition between probability distributions corresponds to the case $\r = (1/n, \dots, 1/n)$. The pre-order that results from relative majorization is also easily checked and again corresponds to checking a small number of inequalities.

It turns out that relative majorization is equally applicable to \emph{quasi-probability} distributions, and therefore can be applied to the Wigner representation of a given fragment. To our knowledge, this application of majorization to quasi-distributions in  physics has not been considered previously. The caveat here is that relative majorization ranges over all stochastic maps with a given fixed point, while in the case of the Wigner representation there is an additional restriction that the phase space structure is respected. Therefore, any relative majorization constraint can provide necessary, but not sufficient conditions on magic state interconversions in the fragment. This observation raises a question that to our knowledge has not been previously considered in the classical statistical mechanics literature: what are the conditions for symplectic majorization of classical statistical mechanics? We shall return to this question later on when we discuss how our approach could be used to construct concrete lower-bounds on distillation rates in~\cref{sec:lower_bounds}.

@@

Given a vector $\x$ we define $\x^\downarrow$ to be the re-arrangement of the components of $\x$ into decreasing order. Now, given two $n$--component vectors $\w$ and $\r$ we define the \emph{Lorenz curve} of $\w$ relative to $\r$, denoted $L_{\w|\r}(x)$, as the piece-wise linear function that passes through $(0,0)$ and the $n$ points
\begin{equation}\label{eq:lorenz}
        (x_k,L_{\w|\r}(x_k)) =\left( \frac{1}{R}\sum_{i=1}^k r_{\pi(i)}, \sum_{i=1}^k w_{\pi(i)} \right),
\end{equation}
where $R:= \sum_{i=1}^n r_i$, and $\pi$ is the permutation on $n$ objects that sends $\widetilde{\w}$ to $\widetilde{\w}^{\downarrow}$, where $\widetilde{w}_i := w_i/r_i$ is the vector of component-wise ratios of $\w$ and $\r$. The form of this requires that $\r$ has no zero components, which we shall assume without loss of generality as in any physical situation the rank of a quantum state is not an operationally meaningful quantity.

If $\w$ and $\r$ are both probability distributions then the Lorenz curve is defined on the interval $[0,1]$, and rises monotonically until it reaches the value $1$ at $x=1$. The value $L_{\w|\r}(x) = 1$ is a global maximum, attained at the end-point $x=1$. However, if $\w$ is a \emph{quasi-probability} distribution with negative values, and $\r$ a regular probability distribution things are different. Now the Lorenz curve is no longer monotone increasing, but is a concave function that breaks through the $L_{\w|\r}(x) = 1$ barrier at an interior point and attains some non-trivial maximum $L_\star$ above the value $1$, before decreasing monotonically to $L_{\w|\r}(x)= 1$ at the end-point $x=1$. See Figure \ref{fig:lcs} for examples of non-monotonic Lorenz curves for quasi-distributions. The non-differentiable points on the curve occur at the locations where $\widetilde{\bmw}$ changes value and are called \emph{elbows}. Within quantum theory, the breaking of this barrier is associated with the degree of non-classicality. Since majorization provides a measure of order/disorder this is consistent with the intuition that negativity, for example within the context of entanglement theory or quantum computing, can provide a greater form of order than is possible within classical theory.

\begin{figure}
    \centering
    \includegraphics[height=4.5cm]{figs/lc_strange.pdf}
    \caption{\textbf{A `heretical' family of Lorenz curves.} Traditionally, Lorenz curves for distributions are monotone increasing cumulant functions that reach a maximum value of $1$. In contrast, Lorenz curves for magic states break through the value of $L(x)=1$ due to the presence of negativity in the associated quasi-probability distributions. The above family of curves correspond to multiple copies of noisy Strange states $\rho_{\rm{S}}(\epsilon)^{\otimes n}$ (\cref{eq:noisysn}) for $n=1,2,3,4$ within a particular stabilizer fragment. Solid lines represent pure Strange states, while dashed lines represent noisy Strange states with depolarising noise ($\epsilon = 0.1$).
    }
    \label{fig:lcs}
\end{figure}

Given the definition of Lorenz curves, the condition for relative majorization can now be phrased compactly. We have that $\w_1 \succ_{\r} \w_2$ if and only if $L_{\w_1 | \r} (x) \ge L_{\w_2 | \r}(x)$ for all $x \in [0,1]$ (technical discussion and proof can be found in~\cref{app:major}. In other words, the Lorenz curve of $\w_1$ relative to $\r$ is never below the Lorenz curve of $\w_2$ relative to $\r$. Phrased more abstractly, the relative majorization pre-order corresponds to the standard ordering of concave, piece-wise linear functions passing through $(0,0)$ and $(1,1)$. We note that another neat geometric formulation of relative majorization exists, stating that the $L_1$-norm angle between $\w_1$ and $\r$ is greater than or equal to the $L_1$-norm angle between $\w_2$ and $\r$ (see \ddd{[CITE]}\nick{haven't seen this before} for details).

@@

If we consider some $\sigma \in \F$, and its corresponding fragment $\R_\sigma$, then we are typically interested in the ability to transform many copies of some magic state $\rho$ into more pure forms of magic. The state $\rho$ is assumed to have negativity in the Wigner representation, so $W_\rho(\x) < 0$ for some regions of $\x \in \P_d$. We again note that we restrict to odd-dimensional quantum systems, for simplicity. In contrast, the state $\sigma$ will have a Wigner distribution $W_\sigma(\x)$ that forms a genuine probability distribution. On the technical assumption that $\sigma$ is full rank, we also have that $W_\sigma(\x) > 0$ for all $\x \in \P_d$. Any non-full rank states $\sigma$ can be handled as a limiting case in which we inject an infinitesimal fraction of depolarising noise $\epsilon (\id/d)$.

As discussed already, the free operations within the magic theory $\R$ are represented by stochastic maps, and within $\R_\sigma$ by stochastic maps that leave $W_\sigma(\x)$ invariant. Therefore, a necessary condition for magic state transformations $\rho_1 \rightarrow \rho_2$ within $\R_\sigma$ will be that 
\begin{equation}
W_{\rho_1} \succ_{W_{\sigma}} W_{\rho_2},
\end{equation}
or put another way, relative to $W_\sigma$ the quasi-distribution $W_{\rho_1}$ is more ordered than the quasi-distribution $W_{\rho_2}$.

This in turn can be expressed in terms of Lorenz curves. To simplify notation we denote by $L_{\rho | \sigma}(x)$ the Lorenz curve $L_{W_{\rho} | W_{\sigma}} (x)$. We therefore have that
\begin{equation}
\rho_1 \rightarrow \rho_2 \mbox{ within }\R_\sigma \mbox{ implies } L_{\rho_1 |\sigma} (x) \ge L_{\rho_2 |\sigma} (x),
\end{equation}
for all $x \in [0,1]$. Therefore, the magic state Lorenz curve condition can be used to impose upper bounds on magic distillation rates. Note, however, that unlike any monotone, this is not a single numerical constraint, but a \emph{family} of constraints. Indeed, for $n$ copies of a $d$--dimensional system the number of terms in $W_{\rho}$ is $d^{2n}$, so the Lorenz curve condition corresponds to \emph{exponentially} many constraints, of which there are as many independent constraints as there are elbows in the Lorenz curve of the target state as shown in~\cref{thm:elbows} in~\cref{app:frag}.
In principle, any one location $x$ provides a valid constraint leading to some upper distillation bound and optimising over the location provides the strictest bound.
